---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load required libraries
library(dplyr)
library(ggplot2)
```


## Loading and preprocessing the data

```{r}
data <- read.table(unz("activity.zip", "activity.csv"), header=T, quote="\"", sep=",")

# Calc avg steps per day
avgByInterval<-aggregate(list(steps=data$steps), by=list(interval=data$interval),  FUN=mean, na.rm=TRUE, na.action=NULL)

```



## What is mean total number of steps taken per day?


```{r}

totalByDay<-aggregate(list(steps=data$steps), by=list(date=data$date),  FUN=sum, na.rm=TRUE, na.action=NULL)

meanSteps<-mean(totalByDay$steps,na.rm=TRUE)
medianSteps<-median(totalByDay$steps,na.rm=TRUE)

hist(totalByDay$steps)

```

- Mean of total steps per day is `r format(meanSteps, digits=2)`.  
- Median total steps per day is `r format(medianSteps, digits=2)`.


## What is the average daily activity pattern?
```{r}
max<-avgByInterval[which.max(avgByInterval$steps),]$interval

plot(steps ~ interval, avgByInterval, type="l")
abline(v=max)

max
```

Max average numbner of steps occurs at interval `r max`  
This represents peak activity at 8:35am

## Imputing missing values
```{r}
missing<-sum(is.na(data$steps))

uniqueDates<-unique(data$date)
uniqueIntervals<-unique(data$interval)

dataImputed<-data

# Impute using daily average
dataImputed$steps[is.na(dataImputed$steps)]<- with(dataImputed, ave(steps, uniqueIntervals, FUN = function(x) mean(x, na.rm = TRUE)))[is.na(dataImputed$steps)]

totalImputedByDay<-aggregate(list(steps=dataImputed$steps), by=list(date=dataImputed$date),  FUN=sum, na.rm=TRUE, na.action=NULL)

hist(totalImputedByDay$steps)

imputedMean<-mean(totalImputedByDay$steps,na.rm=TRUE)
imputedMedian<-median(totalImputedByDay$steps,na.rm=TRUE)

```

There are `r missing` missing values.

Missing values have been imputed by using the average number of steps for the same interval across all days.

After imputation:
- Mean is `r format(imputedMean,digits=2)`
- Median is `r format(imputedMedian,digits=2)`

Occurrences of steps at 5000 have reduced

Both mean and median have increased

## Are there differences in activity patterns between weekdays and weekends?

```{r}
dataImputed$weekpart <- ifelse(weekdays(as.Date(totalImputedByDay$date))%in%c("Saturday","Sunday"),"weekend", "weekday")

dataImputed$weekpart <- as.factor(dataImputed$weekpart)

totalImputedByInterval <- aggregate(list(steps=dataImputed$steps), by=list(interval=dataImputed$interval, weekpart=dataImputed$weekpart),  FUN=mean, na.rm=TRUE, na.action=NULL)

qplot(interval, steps, data=totalImputedByInterval, group=1, facets = .~weekpart,geom="line")
```

On weekdays and weekends there is a spike in activity in the morning.
On the weekend there is more activity later in the day



